---
title: "Covid19 vaccine study Fall2021"
author: "Marcia Alarcon"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 5

---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
TheData <-read.csv(file="fall21.csv",header=TRUE)
TheData <- TheData %>% mutate(across(where(is.character),as.factor))
TheData$treatment <- relevel(TheData$treatment, ref = 'placebo')
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Part of the information you need to compute is the so-called effectiveness of the drug for each subgroup of people. Include a 95% confidence interval of the effectiveness for each different subgroup. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Males

```{r}
malesDataSet = filter(TheData, sex == "guy")
```


```{r}
barchartGC(~infected + treatment,data=malesDataSet)
barchartGC(~infected + treatment,data=malesDataSet, type="percent")

```

The data shows that majority of the population of males were not infected, however the guys who received the treatment actually were more likely to contract Covid than those who received the placebo.

### Numerical Summaries  

```{r}
tableMales <- xtabs(~infected + treatment, data=malesDataSet)
rowPerc(tableMales)
colPerc(tableMales)
```

The numerical summary explains that the majority of the males were okay and uninfected. 41.84% received the placebo while 58.16% received the drug.


## Inferential  

```{r}
chisq.test(tableMales)
chisqtestGC(tableMales)
fisher.test(tableMales) -> fisherResult
print(fisherResult)
```

The Pearson Chi Square test the p-value is 0.0006181

In the Fisher test the p-value is 0.0005801

Get the effectiveness for the odds ratio `r round(fisherResult$estimate,3)`

```{r}
effectiveness = fisherResult$estimate/(fisherResult$estimate+1)*100

```

The drug effectiveness for males is `r effectiveness`%


# Females

```{r}
femalesDataSet = filter(TheData, sex == "gal")
```

```{r}
barchartGC(~infected + treatment,data=femalesDataSet)
barchartGC(~infected + treatment,data=femalesDataSet, type="percent")

```

The data shows that majority of the population of females were not infected, however the females who received the treatment actually were less likely to contract Covid than those who received the placebo.

### Numerical Summaries  

```{r}
tableFemales <- xtabs(~infected + treatment, data=femalesDataSet)
rowPerc(tableFemales)
colPerc(tableFemales)
```

The numerical summary explains that the majority of the females were okay and uninfected. 66.89% received the placebo while 33.11% received the drug.

## Inferential  

```{r}
chisq.test(tableFemales)
chisqtestGC(tableFemales)
fisher.test(tableFemales) -> fisherResult
print(fisherResult)
```

The Pearson Chi Square test the p-value is 0.000000000000003626 which is extremely low.

In the Fisher test the p-value is 0.00000000000000022

Get the effectiveness for the odds ratio `r round(fisherResult$estimate,3)`

```{r}
effectiveness = fisherResult$estimate/(fisherResult$estimate+1)*100

```

The drug effectiveness for females is `r effectiveness`%


# Drug Users

```{r}
DrugUserYesDataSet = filter(TheData, DrugUser == "yes")
```


```{r}
barchartGC(~infected + treatment,data=DrugUserYesDataSet)
barchartGC(~infected + treatment,data=DrugUserYesDataSet, type="percent")

```

The data shows that majority of the population of drug users were not infected, however the drug users who received the treatment actually were less likely to contract Covid than those who received the placebo.

### Numerical Summaries  

```{r}
tableDrugUserYes <- xtabs(~infected + treatment, data=DrugUserYesDataSet)
rowPerc(tableDrugUserYes)
colPerc(tableDrugUserYes)
```

The numerical summary explains that the majority of drug users were okay and uninfected. 66.67% received the placebo while 33.33% received the drug.


## Inferential  

```{r}
chisq.test(tableDrugUserYes)
chisqtestGC(tableDrugUserYes)
fisher.test(tableDrugUserYes) -> fisherResult
print(fisherResult)
```

The Pearson Chi Square test the p-value is 0.003028

In the Fisher test the p-value is 0.002273

Get the effectiveness for the odds ratio `r round(fisherResult$estimate,3)`

```{r}
effectiveness = fisherResult$estimate/(fisherResult$estimate+1)*100

```

The drug effectiveness for drug users is `r effectiveness`%


# LGBTQ

```{r}
LGBTQDataSet = filter(TheData, LGBTQ == "gay")
```


```{r}
barchartGC(~infected + treatment,data=LGBTQDataSet)
barchartGC(~infected + treatment,data=LGBTQDataSet, type="percent")

```

The data shows that majority of the population of LGBTQ were not infected, however the LGBTQ members who received the treatment actually were more likely to contract Covid than those who received the placebo.

### Numerical Summaries  

```{r}
tableLGBTQ <- xtabs(~infected + treatment, data=LGBTQDataSet)
rowPerc(tableLGBTQ)
colPerc(tableLGBTQ)
```

The numerical summary explains that almost half of the LBGTQ members were okay and uninfected. 29.31% received the placebo while 70.69% received the drug.

## Inferential  

```{r}
chisq.test(tableLGBTQ)
chisqtestGC(tableLGBTQ)
fisher.test(tableLGBTQ) -> fisherResult
print(fisherResult)
```

The Pearson Chi Square test the p-value is 0.0000105

In the Fisher test the p-value is 0.000005741

Get the effectiveness for the odds ratio `r round(fisherResult$estimate,3)`

```{r}
effectiveness = fisherResult$estimate/(fisherResult$estimate+1)*100

```

The drug effectiveness for LBGTQ is `r effectiveness`%

# Conclusions and recommendations

In conclusion the men who received the drug were more likely to get Covid than the ones who got the placebo. Females who received the drug were less likely to get Covid than those who got the placebo. Just like the females the drug users had similar reactions. In the LGBTQ community show that those who received the drug were more likely to get covid than the ones who go the placebo just like the mens results. Recommendation regardless of gender, secual orientation, and if one uses drugs or not one need to take the same precautions because whether or not one recieves treatment there is still possibility of contracting covid.
