---
title: "Covid19 vaccine study Fall2021"
author: "Marcia Alarcon"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_depth: 5

---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
TheData <-read.csv(file="fall21.csv",header=TRUE)
TheData <- TheData %>% mutate(across(where(is.character),as.factor))
TheData$treatment <- relevel(TheData$treatment, ref = 'placebo')
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Part of the information you need to compute is the so-called effectiveness of the drug for each subgroup of people. Include a 95% confidence interval of the effectiveness for each different subgroup. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Males

```{r}
malesDataSet = filter(TheData, sex == "guy")
```


```{r}
barchartGC(~infected + treatment,data=malesDataSet)
barchartGC(~infected + treatment,data=malesDataSet, type="percent")

```

The data shows that majority of the population of males were not infected, however the guys who received the treatment actually were more likely to contract Covid than those who received the placebo.

### Numerical Summaries  

```{r}
tableMales <- xtabs(~infected + treatment, data=malesDataSet)
rowPerc(tableMales)
colPerc(tableMales)
```


## Inferential  

```{r}
chisq.test(tableMales)
chisqtestGC(tableMales)
fisher.test(tableMales) -> fisherResult
print(fisherResult)
```


Get the effectiveness for the odds ratio `r round(fisherResult$estimate,3)`

```{r}
effectiveness = fisherResult$estimate/(fisherResult$estimate+1)*100

```

The drug effectiveness for males is `r effectiveness`%


# Females

```{r}
femalesDataSet = filter(TheData, sex == "gal")
```

```{r}
barchartGC(~infected + treatment,data=femalesDataSet)
barchartGC(~infected + treatment,data=femalesDataSet, type="percent")

```

The data shows that majority of the population of females were not infected, however the females who received the treatment actually were less likely to contract Covid than those who received the placebo.

### Numerical Summaries  

```{r}
tableFemales <- xtabs(~infected + treatment, data=femalesDataSet)
rowPerc(tableFemales)
colPerc(tableFemales)
```



## Inferential  

```{r}
chisq.test(tableFemales)
chisqtestGC(tableFemales)
fisher.test(tableFemales) -> fisherResult
print(fisherResult)
```

Get the effectiveness for the odds ratio `r round(fisherResult$estimate,3)`

```{r}
effectiveness = fisherResult$estimate/(fisherResult$estimate+1)*100

```

The drug effectiveness for females is `r effectiveness`%


# Drug Users

```{r}
DrugUserYesDataSet = filter(TheData, DrugUser == "yes")
```


```{r}
barchartGC(~infected + treatment,data=DrugUserYesDataSet)
barchartGC(~infected + treatment,data=DrugUserYesDataSet, type="percent")

```

The data shows that majority of the population of drug users were not infected, however the drug users who received the treatment actually were less likely to contract Covid than those who received the placebo.

### Numerical Summaries  

```{r}
tableDrugUserYes <- xtabs(~infected + treatment, data=DrugUserYesDataSet)
rowPerc(tableDrugUserYes)
colPerc(tableDrugUserYes)
```


## Inferential  

```{r}
chisq.test(tableDrugUserYes)
chisqtestGC(tableDrugUserYes)
fisher.test(tableDrugUserYes) -> fisherResult
print(fisherResult)
```

Get the effectiveness for the odds ratio `r round(fisherResult$estimate,3)`

```{r}
effectiveness = fisherResult$estimate/(fisherResult$estimate+1)*100

```

The drug effectiveness for drug users is `r effectiveness`%



# LGBTQ

```{r}
LGBTQDataSet = filter(TheData, LGBTQ == "gay")
```


```{r}
barchartGC(~infected + treatment,data=LGBTQDataSet)
barchartGC(~infected + treatment,data=LGBTQDataSet, type="percent")

```

The data shows that majority of the population of LGBTQ were not infected, however the LGBTQ members who received the treatment actually were more likely to contract Covid than those who received the placebo.

### Numerical Summaries  

```{r}
tableLGBTQ <- xtabs(~infected + treatment, data=LGBTQDataSet)
rowPerc(tableLGBTQ)
colPerc(tableLGBTQ)
```

## Inferential  

```{r}
chisq.test(tableLGBTQ)
chisqtestGC(tableLGBTQ)
fisher.test(tableLGBTQ) -> fisherResult
print(fisherResult)
```

Get the effectiveness for the odds ratio `r round(fisherResult$estimate,3)`

```{r}
effectiveness = fisherResult$estimate/(fisherResult$estimate+1)*100

```

The drug effectiveness for LBGTQ is `r effectiveness`%

# Conclusions and recommendations

